\chapter{References}
\markboth{REFERENCES}{REFERENCES}
\begin{description}
\setlength{\itemsep}{0pt}
%\setlength{\parsep}{0pt}
%\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt}
\footnotesize

\item [Alley]R.B. 2000. Ice-core evidence of abrupt climate change.
\textit{Proceedings of National Academy of Sciences} 97 (4): 1331--1334.

\item[Apel]J.R. 1987. \textit{Principles of Ocean Physics}.  New York: Academic
Press.

\item [Anderson]J.D. 2005. Ludwig Prandtl's Boundary Layer. \textit{Physics Today} 58 (12): 42--48.

\item [Andersen]O.B., P.L. Woodworth, and R.A. Flather. 1995. Intercomparison
of recent ocean tide models. \textit{J. Geophysical Research} 100 (C12):
25,262--25,282.

\item [Arthur]R.S. 1960. A review of the calculation of ocean currents at the
equator. \textit{Deep-Sea Research} 6 (4): 287--297.

\item [Atlas]R., R.N. Hoffman, and S.C. Bloom. 1993. Surface wind velocity over the oceans. In: \textit{Atlas of satellite observations related to global change}. Edited by R. J. Gurney, J. L. Foster and C. L. Parkinson. 129--140. Cambridge: University Press.

\item [Baker]D.J. 1981. Ocean instruments and experiment design. In
\textit{Evolution of Physical Oceanography: Scientific Surveys in Honor of Henry
Stommel}. Edited by B. A. Warren and C. Wunsch. 396--433. Cambridge: Massachusetts Institute of Technology Press.

\item [Barnett]T.P., M. Latif, N.E. Graham, M. Flugel, S. Pazan, and W. White.
1993. \textsc{enso} and \textsc{enso}-related predictability. Part I: Prediction
of equatorial Pacific sea surface temperature in a hybrid coupled
ocean-atmosphere model. \textit{Journal of Climate} 6: 1545--1566.

\item [Barnier]B., L. Siefridt, and P. Marchesiello. 1995. Thermal forcing for
a global ocean circulation model using a three-year climatology of \textsc{ecmwf}
analyses. \textit{Journal of Marine Systems} 6: 393--380.

\item [Barnston] A.G., Y. Hea, and M.H. Glantz. 1999. Predictive Skill of Statistical and Dynamical Climate Models in SST Forecasts during the 1997--98 El Ni\~{n}o Episode and the 1998 La Ni\~{n}a Onset. \textit{Bulletin of the American Meteorological Society} 80 (2): 217--243.

\item [Batchelor]G.K. 1967. \textit{An Introduction to Fluid Dynamics}.
Cambridge: University Press.

\item [Baumgartner]A., and E. Reichel. 1975. \textit{The World Water Balance}. 
New York: Elsevier.

\item[Beardsley]R.C., A.G. Enriquez, C.A. Friehe, and C.A. Alessi. 1997.
Intercomparison of aircraft and buoy measurements of wind speed and wind stress
during \textsc{smile}. \textit{Journal of Atmospheric and Oceanic
Technology} 14: 969--977.

\item [Behringer]D.W., M. Ji and A. Leetmaa. 1998. An improved coupled model
for \textsc{enso} prediction and implications for ocean initialization: Part 1: The ocean
data assimilation. \textit{Monthly Weather Review} 126(4): 1013--1021.

\item [Bennett]A.F.
1992. \textit{Inverse Methods in Physical Oceanography}.  Cambridge: University Press.

\item [Berlinski]D. 1996. The end of materialist science. \textit{Forbes ASAP}
(December 2, 1996): 146--160.

\item [Binder]R.C. 1949. \textit{Fluid Mechanics}. Second ed. New York:
Prentice--Hall.

\item [Bjerknes]J. 1966. The possible response of the atmospheric Hadley
circulation to equatorial anomalies of ocean temperature. \textit{Tellus} 4:
820-929.

\item [Bjerknes]J. 1972. Large-scale atmospheric response to the 1964--65
Pacific equatorial warming. \textit{Journal of Physical Oceanography} 2 (3):
212--217.

\item [Bjerknes]V. and J.W. Sandstr\"{o}m. 1910. Dynamic Meteorology and
Hydrography, Part I. Statics. Carnegie Institution of Washington DC, Publication
No. 88.

\item [Bleck]R. 2002. An oceanic general circulation model framed in hybrid isopycnic-Cartesian coordinates. \textit{Ocean Modeling} 4: 55--88.

\item [Blumberg]A.F., and G.L. Mellor. 1987. A description of a
three-dimensional ocean circulation model. In: \textit{Three-Dimensional Coastal
Ocean Models}. Edited by N. S. Heaps. 1--16. Washington, DC: American Geophysical
Union.

\item [Bond]G., H. Heinrich, W. Broecker, L. Labeyrie, J. McManus, J. Andrews,
S. Huon, R. Jantschik, S. Clasen, C. Simet, K. Tedesco, M. Klas, G. Bonani, and
S. Ivy. 1992. Evidence for massive discharges of icebergs into the North Atlantic
ocean during the last glacial period. \textit{Nature} 360, 245--.

\item [Boville]B.A., and P.R. Gent. 1998. The \textsc{ncar} Climate System Model, Version
One. \textit{Journal of Climate} 11 (6): 1115--1130.

\item[Bowden]K.F. 1962. Turbulence. In: \textit{The Sea Volume 1}. Edited by M. N.
Hill. 802--825. New York: Interscience Publishers, John Wiley and Sons.

\item [Bracewell]R.N. 1986. \textit{The Fourier Transform and Its Applications}.
Second, revised ed.  New York: McGraw-Hill Publishing Company.

\item [Brauer]A., G. H. Haug, et al. 2008. An abrupt wind shift in western Europe at the onset of the Younger Dryas cold period. \textit{Nature Geoscience} 1 (8): 520--523.
	
\item [Broecker]W.S. 1987. Unpleasant surprises in the greenhouse? \textit{Nature} 328: 123--126.

\item [Broecker]W.S. 1997. Thermohaline circulation, the Achilles heel of our
climate system: Will man-made CO$_2$ upset the current balance? \textit{Science}
278 (5343): 1582--1588.

\item [Bryan] K. 1969. A numerical method for the study of the world ocean.
\textit{Journal of Computational Physics} 4: 347--376.

\item [Bryden]H.L. and T.H. Kinder. 1991. Steady two-layer exchange through the Strait of Gibraltar. \textit{Deep Sea Research} 38 Supplement 1A: S445--S463.
	
\item [Carritt]D.E., and J.H. Carpenter. 1959. The composition of sea water and
the salinity-chlorinity-density problems. National Academy of Sciences-National
Research Council, Report 600: 67--86.

\item [Cartwright]D.E. 1999. \textit{Tides: A Scientific History}. Cambridge,
University Press.
	
\item [Cazenave]A., and J.Y. Royer. 2001. Applications to marine Geophysics. In
\textit{Satellite Altimetry and earth Sciences}. 407--439. San Diego: Academic
Press.

\item [Cess]R.D., M.H. Zhang, P.Minnis, L.Corsetti, E.G. Dutton, B.W. Forgan,
D.P. Garber, W.L. Gates, J.J. Hack, E.F. Harrison, X. Jing, J.T. Kiehl, C.N.
Long, J.-J. Morcrette, G.L. Potter, V. Ramanathan, B. Subasilar, C.H. Whitlock,
D.F. Young, and Y. Zhou. 1995. Absorption of solar radiation by clouds:
Observations versus models. \textit{Science} 267 (5197): 496--499.

\item[Chambers]D.P., B.D. Tapley, and Stewart, R.H. 1998. Measuring heat storage changes in the equatorial Pacific: A comparison between \textsc{Topex} altimetry and Tropical Atmo\-sphere-Ocean buoys. \textit{Journal of Geophysical Research} 103(C9): 18,591--18,597.

\item [Charnock]H. 1955. Wind stress on a water surface. \textit{Quarterly
Journal Royal Meteorological Society} 81: 639--640.

\item [Chelton]D.B., J.C. Ries, B.J. Haines, L.L. Fu, and P.S. Callahan. 2001. Satellite Altimetry. In: \textit{Satellite Altimetry and Earth Sciences: A handbook of techniques and applications}. Editors: L.-L. Fu and A. Cazenave. Academic Press: 1--131.
	
\item [Chen]D., S.E. Zebiak, A.J. Busalacchi, and M.A. Cane. 1995. An improved
proceedure for El Ni\~{n}o forecasting: Implications for predictability.
\textit{Science} 269 (5231): 1699--1702.

\item [Chereskin]T.K., and D. Roemmich. 1991. A comparison of measured and
wind-derived Ekman transport at 11\degrees N in the Atlantic Ocean.
\textit{Journal of Physical Oceanography} 21 (6): 869--878.

\item [Chen]D., S.E. Zebiak, A.J. Busalacchi, and M.A. Cane. 1995. An
improved proceedure for El Ni\~{n}o forecasting: Implications for predictability.
\textit{Science} 269 (5231): 1699--1702.

\item [Chou]S.-H., E. Nelkin, et al. 2004. A comparison of latent heat fluxes over global oceans for four flux products. \textit{Journal of Climate} 17 (20): 3973--3989.

\item [Church]J. A. 2007. Oceans: A Change in Circulation? \textit{Science} 317 (5840): 908--909.
	
\item [Clarke]G.L., G.C. Ewing, and C.J. Lorenzen. 1970. Spectra of
backscattered light from the sea obtained from aircraft as a measure of
chlorophyll concentration. \textit{Science} 167: 1119--1121.

\item [Coakley]J.A., and F.P. Bretherton. 1982. Cloud cover from high
resolution scanner data: Detecting and allowing for partially filled fields of
view. \textit{Journal of Geophysical. Research} 87 (C7): 4917--4932.

\item [Cooley]J.W., P.A. Lewis, and P.D. Welch. 1970. The fast Fourier
transform algorithm: Programming considerations in the calculation of sine,
cosine and Laplace transforms. \textit{Journal of Sound and Vibration} 12:
315--337.

\item [Couper]A. Editor. 1983. \textit{The Times Atlas of the Oceans}. New York:
Van Nostrand Reinhold Company.

\item [Cox]M.D. 1975. A baroclinic model of the world ocean: Preliminary
results. In: \textit{Numerical Models of Ocean Circulation}: 107--120. Washington,
DC: National Academy of Sciences.

\item [Cromwell]T., R.B. Montgomery, and E.D. Stroup. 1954. Equatorial
undercurrent in Pacific Ocean revealed by new methods. \textit{Science} 119
(3097): 648--649.

\item[Cushman-Roisin]B. 1994. \textit{Introduction to Geophysical Fluid
Dynamics}.  Englewood Cliffs: Prentice Hall.

\item[Cutchin]D.L., and R.L. Smith. 1973. Continental shelf waves:
Low-frequency variations in sea level and currents over the Oregon continental
shelf. \textit{Journal of Physical Oceanography} 3 (1): 73--82.

\item [Daley]R. 1991. \textit{Atmospheric Data Analysis}.  Cambridge: University Press.

\item [Danabasoglu]G., J.C. McWilliams, and P.R. Gent. 1994. The role of
mesoscale tracer transports in the global ocean circulation. \textit{Science} 264
(5162): 1123--1126.

\item [Dansgaard]W., S.J. Johnsen, H.B. Clausen, D. Dahl-Johnsen, N.
Gunderstrup, C.U. Hammer, C. Hvidberg, J. Steffensen, A. Sveinbjörnsobttir, J.
Jouze and G. Bond. 1993. Evidence for general instability of past climate from
a 250-kyr ice core record. \textit{Nature} 364: 218-220.

\item [Darnell]W.L., F. Staylor, S.K. Gupta, and F,M. Denn. 1988. Estimation of
surface insolation using sun-synchronous satellite data. \textit{Journal of Climate} 1
(8): 820--835.

\item [Darnell]W.L., W.F. Staylor, S.K. Gupta, N.A. Ritchey, and A.C. Wilbur.
1992. Seasonal variation of surface radiation budget derived from International
Satellite Cloud Climatology Project C1 data. \textit{Journal of Geophysical
Research} 97: 15,741--15,760.

\item [Darwin]Sir G.H. 1911. \textit{The Tides and Kindred Phenomena in the Solar
System}. 3rd ed.  London: John Murray.

\item [DaSilva]A., C.C. Young, and S. Levitus. 1995. Atlas of surface marine
data 1994. Vol. 1: Algorithms and procedures. National Oceanic and Atmospheric
Administration Report.

\item [Davis]R.A. 1987. \textit{Oceanography: An Introduction to the Marine
Environment}. Dubuque: Wm. C. Brown Publishers.

\item [Davis]R.E., R. DeSzoeke, and P. Niiler. 1981. Variability in the upper
ocean during \textsc{mile}. Part II: Modeling the mixed layer response. \textit{Deep-Sea
Research} 28A (12): 1453--1475.

\item [Davis]R.E., D.C. Webb, L.A. Regier, and J. Dufour. 1992. The
Autonomous Lagrangian Circulation Explorer (\textsc{alace}). \textit{Journal of Atmospheric and Oceanic Technology} 9: 264--285.

\item [Defant]A. 1961. \textit{Physical Oceanography}. New York: Macmillan
Company.

\item [Dietrich]G., K. Kalle, W. Krauss, and G. Siedler. 1980. \textit{General
Oceanography}. 2nd ed. Translated by Susanne and Hans Ulrich Roll. New York: John
Wiley and Sons (Wiley-Interscience).

\item [Dittmar]W. 1884. Report on researches into the composition of ocean
water, collected by the HMS Challenger. \textit{Challenger Reports, Physics and
Chemistry} 1.

\item [Dobson]G.M.B. 1914. Pilot balloon ascents at the central flying school,
Upavon, during the year 1913. \textit{Quarterly Journal of the Royal
Meteorological Society} 40: 123--135.

\item [Domingues]C. M., J. A. Church, et al 2008. Improved estimates of upper-ocean warming and multi-decadal sea-level rise. \textit{Nature} 453 (7198): 1090--1093.
	
\item [Doodson]A.T. 1922. Harmonic development of the
tide-generating potential. \textit{Proceedings of the Royal Society of London} A
100: 305--329.

\item [Dorman]C.E. and R.H. Bourke. 1981. Precipitation over the Atlantic
ocean, 30\degrees S to 70\degrees N. \textit{Monthly Weather Review} 109:
554--563.

\item [Dritschel]D.G., M. de la T. Juarez and M.H.P. Ambaum. 1999. The
three-dimensional vortical nature of atmospheric and oceanic turbulent flows.
\textit{Physics of Fluids} 11(6): 1512--1520.

\item [Dushaw]B.D., P.F. Worcester, B.D. Cornuelle, and B.M. Howe. 1993. On
equations for the speed of sound in sea water.
\textit{Journal of the Acoustical Society of America} 93: 255--275.

\item [Ebbesmeyer]C.C., and W.J. Ingraham. 1992. Shoe spill in the North
Pacific. \textit{EOS, Transactions of the American Geophysical Union} 73 (34):
361, 365.

\item [Ebbesmeyer]C.C., and W.J. Ingraham. 1994. Pacific toy spill fuels ocean
current pathways research. \textit{EOS Transactions of the American Geophysical
Union} 75 (37): 425, 427, 430.

\item [Eden]C., and J. Willebrand. 1999. Neutral density revisited.
\textit{Deep-Sea Research} Part II: Topical Studies in Oceanography. 46: 33--54.

\item [Egbert]G.B. and R.D. Ray 2000. Significant dissipation of tidal energy in the deep ocean inferred from satellit altimeter data. \textit{Nature} 405: 775--778.

\item [Ekman]V.W. 1905. On the influence of the Earth's rotation on ocean
currents. \textit{Arkiv for Matematik, Astronomi, och Fysik}: 2 (11).

\item [Emery]W., and P. Schussel. 1989. Global difference between skin and bulk
sea surface temperatures. \textit{EOS: Transactions of the American Geophysical
Union} 70 (14): 210--211.

\item [Feynman]R.P., R.B. Leighton, and M. Sands. 1964. \textit{The Feynman
Lectures on Physics}.  Addison-Wesley Publishing Company.

\item [Fofonoff]N.P., and R.C. Millard. 1983. Algorithms for computation of
fundamental properties of sea water. \textsc{unesco} Technical Papers in Marine
Science 44.

\item [Friedman]R.M. 1989. \textit{Appropriating the Weather. Vilhelm Bjerknes
and the Construction of a Modern Meteorology.}  Ithaca and London: Cornell
University Press.

\item [Friedrichs]M.A.M., and M.M. Hall. 1993. Deep circulation in the tropical
North Atlantic. \textit{Journal of Marine Research} 51 (4): 697--736.

\item[Freilich]M.H., and R.S. Dunbar. 1999. The accuracy of the NSCAT 1 vector
winds: Comparisons with National Data Buoy Center buoys. \textit{Journal of
Geophysical Research} submitted 

\item [Garabato]A.C.N., K.L. Polzin, B.A. King, K.J. Heywood, and M. Visbeck. 2004. Widespread Intense Turbulent Mixing in the Southern Ocean. \textit{Science} 303 (5655): 210--213.

\item [Garabato]A.C.N., D.P. Stevens, A.J. Watson, and W. Roether. 2007. Short-circuiting of the overturning circulation in the Antarctic Circumpolar Current. \textit{Nature} 447 (7141): 194--197.
	
\item[Gargett]A.E., and G. Holloway. 1992. Sensitivity of the \textsc{gfdl}
ocean model to different diffusivities of heat and salt. \textit{Journal of
Physical Oceanography} 22 (10): 1158--1177.

\item [Garrett]C. 2006. Turbulent dispersion in the ocean. \textit{Progress in Oceanography} 70 (2--4): 113--125.

\item[Gates]W.L., A. Henderson-Sellers, G.J. Boer, C.K. Folland, A. Kitoh, B.J. McAvaney, F. Semazzi, N. Smith, A.J. Weaver, and Q.-C. Zeng. 1996. \textit{Climate Models--Evaluation}. In: \textit{Climate Change 1995}. Edited by J.T. Houghton, L.G.M. Filho, B.A. Callander, N. Harris, A. Kattenberg and K.
Maskell. 229--284. Cambridge: University Press.

\item [Gates]W.L. 1992. \textsc{amip}: The Atmospheric Model Intercomparison Project.
\textit{Bulletin American Meteorological Society} 73: 1962--1970.

\item[Gent]P.R., and J.C. McWilliams. 1990. Isopycnal mixing in ocean
circulation models. \textit{Journal of Physical Oceanography} 20: 150--155.

\item [Gleckler]P.J., and B. Weare. 1997. Uncertainties in global ocean surface
heat flux climatologies derived from ship observance. \textit{Journal of Climate}
10: 2763--2781.

\item [Gill]A.E. 1982. \textit{Atmosphere-Ocean Dynamics}. New York: Academic Press.

\item [Gnadadesikan]A. 1999. A simple predictive model for the structure of the
oceanic pycnocline. \textit{Science} 283 (5410): 2077--2079.

\item [Goldenberg]S.B., and J.J. O'Brien. 1981. \textit{Monthly Weather Review}
109: 1190.

\item [Goldstein]S. 1965. \textit{Modern Developments in Fluid Dynamics}: Two
Volumes. New York: Dover Publications.

\item [Gordon]H.R., D.K. Clark, J.W. Brown, O.B. Brown, R.H. Evans, and W.W.
Broenkow. 1983. Phytoplankton pigment concentrations in the Middle Atlantic
Bight: comparison of ship determinations and \textsc{czcs} estimates. \textit{Applied 
Physics} 22 (1): 20-36.

\item [Gouretski]V., and K. Jancke. 1995. A consistent pre-\textsc{woce}
hydrographic data set for the south Atlantic: Station data and gridded fields.
\textsc{woce} Report No. 127/95.

\item [Graber]H.C., V.J. Cardone, R.E. Jensen, D.N. Slinn, S.C. Hagen, A.T. Cox, M.D. Powell, and 
C. Grassl. 2006. Coastal Forecasts and Storm Surge Predictions for Tropical Cyclones: A Timely Partnership Program. \textit{Oceanography} 19 (1): 131--141.
	
\item [Grassl]H. 2000. Status and improvements of coupled general circulation
models. \textit{Science} 288 (5473): 1991--1997.

\item[Gregg]M.C. 1987. Diapycnal mixing in the thermocline: A review.
\textit{Journal of Geophysical Research} 92 (C5): 5,249--5,289.

\item [Gregg]M.C. 1991. The study of mixing in the ocean: a brief history.
\textit{Oceanography} 4 (1): 39--45.

\item [Hackett]B., L.P. Roed, B. Gjevik, E.A. Martinsen, and L.I. Eide. 1995. A
review of the Metocean Modeling Project (\textsc{mompop}) Part 2: Model
validation study. In: \textit{Quantitative Skill Assessment for Coastal Ocean
Models}. Edited by D. R. Lynch and A. M. Davies. 307--327. Washington DC:
American Geophysical Union.

\item [Haidvogel]D.B., and A. Beckmann. 1998. Numerical models of the coastal ocean. In: \textit{The Sea, Volume 10}. Edited by K. H. Brink and A. R. Robinson. 457--482. New York: John Wiley and Sons.

\item [Haidvogel]D.B. and A. Beckmann. 1999. \textit{Numerical Ocean Circulation
Modeling}. London, Imperial College Press.

\item [Hall]M.M., and H.L. Bryden. 1982. Direct
estimates and mechanisms of ocean heat transport. \textit{Deep-Sea Research} 29:
339--359.

\item [Harrison]D.E. 1989. On climatological monthly mean wind stress and wind
stress curl fields over the world ocean. \textit{Journal of Climate} 2: 57.

\item [Harrison]D.E., and N.K. Larkin. 1996. The \textsc{coads} sea level pressure
signal: A near-global el Ni\~{n}o composite and time series view, 1946--1993.
\textit{Journal of Climate} 9 (12): 3025--3055.

\item [Harrison]D.E. and N.K. Larkin 1998. El Ni\~{n}o-Southern Oscillation sea surface temperature and wind anomalies, 1946--1993. \textit{Reviews of Geophysics} 36 (3): 353--399.

\item [Hartmann]D.L. 1994. \textit{Global Physical Climatology}.
Academic Press.

\item[Hasselmann]K. 1961. On the non-linear energy transfer in a gravity-wave
spectrum Part 1. General theory. \textit{Journal of Fluid Mechanics} 12 (4):
481--500.

\item[Hasselmann]K. 1963a. On the non-linear energy transfer in a gravity wave
spectrum. Part 2. Conservation theorems; wave-particle analogy; irreversibility.
\textit{Journal of Fluid Mechanics} 15 (2): 273--281.

\item[Hasselmann]K. 1963b. On the non-linear energy transfer in a gravity wave
spectrum. Part 3. Evaluation of the energy flux and swell-sea interaction for a
Neumann spectrum. \textit{Journal of Fluid Mechanics}. 15 (3): 385--398.

\item[Hasselmann]K. 1966. Feynman diagrams and interaction rules of wave-wave
scattering processes. \textit{Reviews of Geophysical}. 4 (1): 1--32.

\item [Hasselmann]K. 1970. Wind--driven inertial oscillations. \textit{Geophysical Fluid Dynamics} 1: 1 463--502.

\item [Hasselmann]K., T.P. Barnett, E. Bouws, H. Carlson, D.E. Cartwright, K.
Enke, J.A. Ewing, H. Gienapp, D.E. Hasselmann, P. Kruseman, A. Meerburg, P.
Müller, D.J. Olbers, K. Richter, W. Sell, and H. Walden. 1973. Measurements of
wind-wave growth and swell decay during the Joint North Sea Wave Project
(\textsc{jonswap}). \textit{Ergänzungsheft zur Deutschen Hydrographischen
Zeitschrift Reihe} A(8°) (Nr. 12): 95.

\item [Hasselmann]K., and S. Hasselmann. 1991. On the nonlinear mapping of an
ocean wave spectrum into a synthetic aperture radar image spectrum and its
inversion. \textit{Journal of Geophysical Research} C96 10,713--10,729.

\item [Heaps]N.S., Editor. 1987. \textit{Three-Dimensional Coastal Ocean Models}.
Washington DC: American Geophysical Union.

\item [Hinze]J.O. 1975. \textit{Turbulence}. 2nd ed. New York: McGraw-Hill.

\item [Hirst]A.C., S.P. O'Farrell, and H.B. Gordon. 2000. Comparison of a
coupled ocean-atmosphere model with and without oceanic eddy-induced advection.
Part I: Oceanic spinup and control integration. \textit{Journal of Climate} 13
(1): 139--163.

\item [Hoffman]D., and O.J. Karst. 1975. The theory of the Rayleigh distribution
and some of its applications. \textit{Journal of Ship Research} 19 (3): 172--191.

\item [Hogg]N., J. McWilliams, P. Niiler and J. Price. 2001. Objective 8---To
determine the important processes and balances for the dynamics of the general
circulation. In: \textit{2001 U.S. WOCE Report}. College Station, Texas, U.S. \textsc{woce}
Office: 50--59.

\item [Holloway]G. 1986. Eddies, waves, circulation, and mixing:
statistical geofluid mechanics. \textit{Annual Reviews of Fluid Mechanics} 18:
91--147.

\item [Holloway]G. 1986. Estimation of oceanic eddy transports from satellite
altimetry. \textit{Nature} 323 (6085): 243--244.

\item [Holloway]G. 1994. Representing eddy forcing in models. \textsc{woce} Notes
6 (3): 7--9.

\item [Horikawa]K. 1988. \textit{Nearshore Dynamics and Coastal Processes}.
Tokyo: University of Tokyo Press.

\item [Houghton]J.T. 1977. \textit{The Physics of Atmospheres}. Cambridge: University Press.

\item [Houghton]J.T., L.G.M. Filho, B.A. Callander, N. Harris, A. Kattenberg,
and K. Maskell. 1996. \textit{Climate Change 1995: The Science of Climate
Change}. Cambridge: University Press.

\item [Hoyt]D.V., and K.H. Schatten. 1997. \textit{The Role of the Sun in Climate
Change}.  Oxford: Oxford University Press.

\item [Huffman]G.J., R.F. Adler, B. Rudolf, U. Schneider, and P.R. Keehm. 1995.
Global precipitation estimates based on a technique for combining
satellite-based estimates, rain gauge analysis, and \textsc{nwp} model
precipitation information. \textit{Journal of Climate} 8: 1284--1295.

\item[Huffman]G.J., R.F. Adler, P.A. Arkin, A. Chang, R. Ferraro, A. Gruber, J.
Janowiak, A. McNab, B. Rudolf, and U. Schneider. 1997. The Global Precipitation
Climatology Project (\textsc{gpcp}) combined precipitation data set.
\textit{Bulletin of the American Meteorological Society} 78 (1): 5--20.

\item [Ichiye]T., and J. Petersen. 1963. The anomalous rainfall of the
1957--1958 winter in the equatorial central Pacific arid area. \textit{Journal of
the Meteorological Society of Japan} Series II, 41: 172--182.

\item [International Hydrographic Bureau]1953. Limits of oceans and seas, 3rd ed. Special Report No. 53, Monte Carlo.

\item [IPCC] Intergovernmental Panel on Climate Change. 2007. \textit{Climate Change 2007: The Physical Science Basis. Contribution of Working Group I to the Fourth Assessment Report of the Intergovernmental Panel on Climate Change}. [Solomon, S., D. Qin, M. Manning, Z. Chen, M. Marquis, K.B. Averyt, M. Tignor and H.L. Miller (eds.)]. Cambridge University Press, Cambridge, United Kingdom and New York, NY, USA, 996 pp. 

\item [Iselin]C. 1936. A study of the circulation of the Western North Atlantic.
\textit{Physical Oceanography and Meteorology}. 6 (4): 101.

\item [Isemer]H.J., and L. Hasse. 1987. \textit{The Bunker Climate Atlas of the
North Atlantic}. Volume 2.  Berlin: Springer-Verlag.

\item [Jackett]D.R., and T.J. McDougall. 1997. A neutral density variable for
the world's oceans. \textit{Journal of Physical Oceanography} 27: 237--263.

\item [Jan van Oldenborgh] G., M.A. Balmaseda, L. Ferranti, T.N. Stockdale, and D.L.T. Anderson. 2005. Did the ECMWF Seasonal Forecast Model Outperform Statistical \textsc{enso} Forecast Models over the Last 15 Years? \textit{Journal of Climate} 18 (16): 3240--3249.

\item [Jarosz]E., D.A. Mitchell, D.W. Wang, and W.J. Teague. 2007. Bottom-Up Determination of Air-Sea Momentum Exchange Under a Major Tropical Cyclone. \textit{Science} 315 (5819): 1707--1709.

\item [Jayne]S.R., L.C.S. Laurent and S.T. Gille. 2004. Connections between ocean bottom topography and Earth's climate. \textit{Oceanography} 17 (1): 65--74.

\item [Jelesnianski]C.P.J., P.C. Chen, and W.A. Shaffer. 1992. \textsc{slosh}: Sea, lake, and overland surges from hurricanes. \textsc{noaa} Technical Report \textsc{nws} 48.

\item[Jerlov]N.G. 1976. \textit{Marine Optics}. Amsterdam: Elsevier Scientific
Publishing Company.

\item[Ji]M., A. Leetmaa, and V. Kousky. 1996. Coupled model predictions of \textsc{enso} during the 1980s and the 1990s at the National Centers for Environmental Prediction. \textit{Journal of Climate} 9 (12): 3105--3120.

\item[Ji]M., D.W. Behringer, and A. Leetmaa. 1998. An improved coupled model for \textsc{enso} prediction and implications for ocean initialization. Part II: The coupled model. \textit{Bulletin of the American Meteorological Society} 126 (4): 1022--1034.

\item [Johns]E., D.R. Watts, and H.T. Rossby. 1989. A test of geostrophy in the
Gulf Stream. \textit{Journal of Geophysical Research} 94 (C3): 3211--3222.

\item [Johns]T.C., R.E. Carnell, J.F. Crossley, J.M. Gregory, J.F.B. Mitchell,
C.A. Senior, S.F.B. Trett, and R.A. Wood. 1997. The second Hadley Centre coupled
ocean-atmosphere GCM: model description, spinup and validation. \textit{Climate
Dynamics} 13 (2): 103--134.

\item[Joseph]J., and H. Sender. 1958. Uber die horizontale diffusion im meere.
\textit{Deutsches Hydrographiches Zeitung} 11: 49--77.

\item [Josey]S.A., E.C. Kent, and P.K. Taylor. 1999. New insights into the ocean heat budget closure problem from analysis of the \textsc{soc} air-sea flux climatology. \textit{Journal of Climate} 12: 2856--2880.
	
\item [JPOTS]Joint Panel on Oceanographic Tables and Standards. 1981. \textit{The practical salinity scale 1978 and the international equation of state of seawater 1980}. Paris: \textsc{unesco} Technical Papers in Marine Science 36: 25.
	
\item [JPOTS]Joint Panel on Oceanographic Tables and Standards. 1991. \textit{Processing of Oceanographic Station Data}. Paris: \textsc{unesco}.

\item [Kalnay]E., M. Kanamitsu, R. Kistler, W. Collins, D. Deaven, L. Gandin,
M. Iredell, S. Saha, G. White, J. Woollen, Y. Zhu, M. Chelliah, W. Ebisuzaki, W.
Higgins, J. Janowiak, K.C. Mo, C. Ropelewski, J. Wang, A. Leetmaa, R. Reynolds,
R. Jenne, and D. Joseph. 1996. The \textsc{ncep/ncar} 40--year reanalysis
project. \textit{Bulletin American Meteorological Society} 77: 437--471.

\item[Kantha]L.H. 1998. Tides--A modern perspective. \textit{Marine Geodesy} 21:
275--297.

\item [Kent]E.C., and P.K. Taylor. 1997. Choice of a Beaufort Scale.
\textit{Journal of Atmospheric and Oceanic Technology} 14 (2): 228--242.

\item[Kerr]R.A. 1998. Models win big in forecasting El Ni\~{n}o. \textit{Science} 280
(5363): 522--523.

\item [Kerr]R.A. 2002. Salt fingers mix the sea. \textit{Science} 295 (5561):
1821.

\item [Kiehl]J.T., and K.E. Trenberth. 1996. Earth's annual global mean energy
budget. \textit{Bulletin of the American Meteorological Society} 78 (2):
197--208.

\item [Kilpatrick]K.A., G.P. Podesta, and R. Evans. 2001. Overview of the \textsc{noaa/nasa} advanced very high resolution radiometer Pathfinder algorithm for sea surface temperature and associated
matchup database. \textit{Journal of Geophysical Research} 106: 9179--9198.

\item [Kistler]R.E., E. Kalnay, W. Collins, S. Saha, G. White, J. Woolen, M.
Chelliah, and W. Ebisuzaki. 2000. The \textsc{ncep/ncar} 50-year reanalysis.
\textit{Bulletin of the American Meteorological Society} 82: 247--267.

\item [Komen]G.J., L. Cavaleri, M. Donelan, K. Hasselmann, S. Hasselmann, and
P.A.E.M. Jans\-sen. 1996. \textit{Dynamics and Modelling of Ocean Waves}. 1st
paperback ed.  Cambridge: University Press.

\item [Kullenberg]B. 1954. Vagn Walfrid Ekman 1874--1954. \textit{Journal du Conseil international pour l'exploration de la mer} 20 (2): 140--143.

\item [Kumar]A., A. Leetmaa, and M. Ji. 1994. Simulations of atmospheric
variability induced by sea surface temperatures and implications for global
warming. \textit{Science} 266 (5185): 632--634.

\item [Kundu]P.K. 1990. \textit{Fluid Mechanics}. San Diego: Academic Press.

\item [Kunze]E., and J.M. Toole. 1997. Tidally driven vorticity, diurnal shear,
and turbulence atop Fieberling Seamount. \textit{Journal of Physical
Oceanography} 27 (2): 2,663--2,693.

\item [Lagerloef]G.S.E., G. Mitchum, R. Lukas, and P. Niiler. 1999. Tropical
Pacific near-surface current estimates from altimeter, wind and drifter data.
\textit{Journal of Geophysical Research} 104 (C10): 23,313--23,326.

\item [Lamb]H. 1945. \textit{Hydrodynamics}. 6th, first American edition. New
York: Dover Publications.

\item [Lang]K.R. 1980. \textit{Astrophysical Formulae: A Compendium for the
Physicist and Astrophysicist}. 2nd ed.  Berlin: Springer-Verlag.

\item [Langer]J. 1999. Computing in physics: Are we taking it too seriously? Or
not seriously enough? \textit{Physics Today} 52 (7): 11--13.

\item [Langmuir]I. 1938. Surface motion of water induced by wind. \textit{Science}
87: 119--123.

\item [Larson]R. 2002. E-Enabled textbooks: Lower cost, higher
functionality. \textit{Syllabus} 15(10): 44.

\item [Latif]M., A. Sterl, E. Maier-Reimer, and M.M. Junge.
1993. Structure and predictability of the El Ni\~{n}o/Southern Oscillation
phenomenon in a coupled ocean-atmosphere model. \textit{Journal of Climate} 6:
700--708.

\item [Lawrence]M.G., J. Landgraf, P. Jöckel, and B. Eaton. 1999. Artifacts in
global atmospheric modeling: Two recent examples. \textit{EOS Transactions
American Geophysical Union} 80 (11): 123, 128.

\item [Lean]J., J. Beer, and R. Bradley. 1995. Reconstruction of solar
irradiance since 1610: Implications for climate change. \textit{Geophysical
Research Letters} 22 (23): 3195--3198.

\item [Ledwell]J.R., A.J. Watson and C.S. Law 1998. Mixing of a tracer in the pycnocline. \textit{Journal of Geophysical Research} 103(C10): 21,499--421,529.

\item[Leetmaa]A., and A.F. Bunker. 1978. Updated charts of the
mean annual wind stress, convergences in the Ekman layers and Sverdrup transports
in the North Atlantic. \textit{Journal of Marine Research} 36: 311--322.

\item[Leetmaa]A., J.P. McCreary, and D.W. Moore. 1981. Equatorial currents;
observation and theory. In: \textit{Evolution of Physical Oceanography}. Edited
by B. A. Warren and C. Wunsch. 184--196. Cambridge: Massachusetts Institute of Technology Press.

\item [LeProvost]C., M.L. Genco, F. Lyard, P. Vincent, and P. Canceil. 1994. Spectroscopy of the world ocean tides from a finite element hydrodynamic model. \textit{Journal Geophysical Research} 99 (C12): 24,777--24,797.

\item [LeProvost]C., A. F. Bennett and D. E. Cartwright. 1995. Ocean tides for and from Topex/ Poseidon. \textit{Science} 267 (5198): 639--647.

\item [Levitus]S. 1982. \textit{Climatological Atlas of the World Ocean}. \textsc{noaa} Professional Paper 13.

\item [Levitus]S. 1994. \textit{World Ocean Atlas 1994 \textsc{cd-rom} Data Set}. \textsc{noaa} National Oceanographic Data Center.

\item [Lewis]E.L. 1980. The Practical Salinity
Scale 1978 and its antecedents.
\textit{IEEE Journal of Oceanic Engineering} \textsc{oe}-5: 3--8.

\item [List]R.J. 1966. \textit{Smithsonian Meteorological Tables}. 6th ed. Washington DC: The Smithsonian Institution.

\item [Liu]W.T. 2002. Progress in scatterometer application. \textit{Journal of Oceanography} 58: 121--136.

\item [Longuet-Higgins]M.S., and O.M. Phillips. 1962. Phase velocity effects in tertiary wave interactions. \textit{Journal of Fluid Mechanics.} 12 (3): 333--336.

\item [Lynch]D.R., J.T.C. Ip, C.E. Naimie, and F.E. Werner. 1996. Comprehensive
coastal circulation model with application to the Gulf of Maine.
\textit{Continental Shelf Research} 16 (7): 875--906.

\item [Lynn]R.J., and J.L. Reid. 1968. Characteristics and circulation of deep
and abyssal waters. \textit{Deep-Sea Research} 15 (5): 577--598.

\item [McAvaney]B.J., C. Covey, S. Joussaume, V. Kattsov, A. Kitoh, W. Ogana,
A.J. Pitman, A.J. Weaver, R. A. Wood and Z.-C. Zhao. 2001. Model Evaluation. In:
\textit{Climate Change 2001: The Scientific Basis. Contribution of Working Group 1
to the Third Assessment Report of the Intergovernmental Panel on Climate Change}.
Editeb by J.T. Houghton, Y. Ding, D.J. Griggs, N. Noguer, P.J. v. d. Linden, X.
Dai, K. Maskell and C.A. Johnson. Cambridge, University Press: 881.

\item [MacKenzie]K.V. 1981. Nine-term equation for sound speed in the ocean.
\textit{Journal of the Acoustical Society of America} 70: 807--812.

\item [McDougall]T.J. 1987. Neutral surfaces. \textit{Journal of Physical Oceanography} 17 (11): 1950--1964.

\item [McDougall]T. J. and R. Feistel 2003. What causes the adiabatic lapse rate? \textit{Deep Sea Research Part I: Oceanographic Research Papers} 50 (12): 1523--1535.

\item [McIntyre]M.E. 1981. On the `wave momentum' myth. \textit{Journal of Fluid Mechanics} 106: 331--347.

\item [McNally]G.J., W.C. Patzert, A.D. Kirwan, and A.C. Vastano. 1983. The
near-surface circulation of the North Pacific using satellite tracked drifting
buoys. Journal of Geophysical Research 88 (C12): 7,507--7,518.

\item [McPhaden]M.J. 1986. The equatorial undercurrent: 100 years of discovery. \textit{EOS Transansactions of the Amererican Geophysical Union} 67 (40): 762--765.

\item [McPhaden]M.J., A.J. Busalacchi, R. Cheney, J.R. Donguy, K.S. Gage, D. Halpern, M. Ji, P. Julian, G. Meyers, G.T. Mitchum, P.P. Niiler, J. Picaut, R.W. Reynolds, N. Smith K. Takeuchi. 1998. The Tropical Ocean-Global Atmosphere (textsc{toa}) observing system: A decade of progress.
\textit{Journal of Geophysical Research} 103: 14,169--14,240.
	
\item [McPhaden]M.J. 1999. Genesis and evolution of the 1997-1998 El Ni\~{n}o. \textit{Science} 283 (5404): 950--954.

\item [McPhaden]M.J., S.E. Zebiak, and M.H. Glantz. 2006. \textsc{enso} as an Integrating Concept in Earth Science. \textit{Science} 314 (5806): 1740--1745.

\item [Ma]X.C., C.K. Shum, R.J. Eanes, and B.D. Tapley. 1994. Determination of ocean tides from the first year of Topex/Poseidon altimeter measurements. \textit{Journal of Geophysical Research} 99 (C12): 24,809--24,820.

\item [Malanotte-Rizzoli]P., Ed. 1996. \textit{Modern Approaches to Data Assimilation in Ocean Modeling}. Amsterdam: Elsevier.

\item [Maltrud]M.E., R.D. Smith, A.J. Semtner, and R.C. Malone. 1998. Global eddy-resolving ocean simulations driven by 1985--1995 atmospheric winds.
\textit{Journal of Geophysical Research} 103 (C13): 30,825--30,852.

\item [Margules]M. 1906. Uber Temperaturschichtung in stationarbewegter und
ruhender Luft. \textit{Meteorologische Zeitschrift} 241--244.

\item [Marotzke]J., and J.R. Scott. 1999. Convective mixing and the
thermohaline circulation. \textit{Journal of Physical Oceanography} 29 (11):
2962--2970.

\item [Marotzke]J. 2000. Abrupt climate change and thermohaline circulation:
Mechanisms and predictability. \textit{Proceedings National Academy of Sciences}
97 (4): 1347--1350.

\item [Martrat]B., J.O. Grimalt, N.J. Shackleton, L. de Abreu, M.A. Hutterli, and T.F. Stocker. 2007. Four Climate Cycles of Recurring Deep and Surface Water Destabilizations on the Iberian Margin. \textit{Science} 317 (5837): 502--507.

\item[Matth\"{a}us]W. 1969. Zur entdeckungsgeschichte des \"{A}quatorialen
Unterstroms im Atkantischen Ozean. \textit{Beitrage Meereskunde} 23: 37--70.

\item [Maury]M.F. 1855. \textit{Physical Geography of the Sea}.  Harper.

\item[May]D.A., M.M. Parmenter, D.S. Olszewski, and B.D. McKenzie. 1998.
Operational processing of satellite sea surface temperature retrievals at the
Naval Oceanographic Office. \textit{Bulletin of the American Meteorological
Society} 79 (3): 397--407.

\item [Mellor]G.L., and
T. Yamada. 1982. Development of a turbulence closure model for geophysical fluid
problems. \textit{Reviews of Geophysics and Space Physics} 20 (4): 851--875.

\item [Mellor]G.L. 1998. \textit{User's Guide for a Three-dimensional, Primitive equation, Numerical Ocean Model Version 1998}. Princeton, Princeton University: 41.
	
\item [Menard]H.W., and S.M. Smith. 1966. Hypsometry of ocean basin provinces.
\textit{Journal of Geophysical Research} 71: 4305--4325.

\item [Mercier]H., M. Arhan and J.R.E. Lutjeharm. 2003. Upper-layer
circulation in the eastern Equatorial and South Atlantic Ocean in January--March
1995. \textit{Deep-Sea Research} 50 (7): 863--887.

\item [Merryfield]W.J., G. Holloway, and A.E. Gargett.
1999. A global ocean model with double-diffusion mixing. \textit{Journal of
Physical Oceanography} 29 (6): 1124--1142.

\item [Miles]J.W. 1957. On the generation of surface waves by shear flows.
\textit{Journal of Fluid Mechanics}. 3 (2) 185--204.

\item [Millero]F.J., G. Perron, and J.F. Desnoyers. 1973. Heat capacity of
seawater solutions from 5\degrees\  to 35\degrees C and 0.05 to
22$^\circ/_{\circ\circ}$  chlorinity. \textit{Journal of Geophysical Research} 78
(21): 4499--4506.

\item[Millero]F.J., C.-T. Chen, A. Bradshaw, and K. Schleicher. 1980. A new
high pressure equation of state for seawater. \textit{Deep-Sea Research} 27A:
255--264.

\item [Millero]F.J., and A. Poisson. 1981. International one-atmosphere
equation of state of seawater.\textit{Deep-Sea Research} 28A (6): 625--629.

\item [Millero]F.J. 1996. \textit{Chemical Oceanography} (2nd ed). New York, CRC
Press.

\item [Millero]F. J., R. Feistel, et al. 2008. The composition of Standard Seawater and the definition of the Reference-Composition Salinity Scale. \textit{Deep Sea Research Part I: Oceanographic Research Papers} 55 (1): 50--72.

\item [Montgomery]R.B., and E.D. Stroup. 1962. Equatorial Waters and Currents at
150\degrees W in July--August, 1952. Baltimore: The Johns Hopkins Press.

\item [Morel]A. 1974. Optical porperties of pure water and pure seawater. In:
\textit{Optical Aspects of Ocean\-ography}. Edited by N. G. Jerlov and E. S.
Nielson.  1--24. Academic Press.
	
\item [Moskowitz]L. 1964. Estimates of the power spectrums for fully developed
seas for wind speeds of 20 to 40 knots. \textit{Journal of Geophysical Research} 69 (24):
5161--5179.

\item [Moum]J.N., and D.R. Caldwell. 1985. Local influences on shear-flow
turbulence in the equatorial ocean. \textit{Science} 230: 315--316.

\item [Munk]W.H. 1950. On the wind-driven ocean circulation. \textit{Journal of Meteorology} 7 (2): 79--93.

\item [Munk]W.H., and E. Palmen. 1951. A note on the dynamics of the Antarctic
Circumpolar Current. \textit{Tellus} 3: 53--55.

\item [Munk]W.H. 1966. Abyssal recipes. \textit{Deep-Sea Research} 13: 707--730.

\item [Munk]W.H., G.R. Miller, F.E. Snodgrass, and N.F. Barber. 1963.
Directional recording of swell from distant storms. \textit{Philosophical
Transactions Royal Society of London} 255 (1062): 505--584.

\item [Munk]W.H., and D.E. Cartwright. 1966. Tidal spectroscopy and prediction.
\textit{Philosophical Transactions Royal Society London} Series A. 259 (1105):
533--581.

\item [Munk]W.H., R.C. Spindel, A. Baggeroer, and T.G. Birdsall. 1994. The
Heard Island feasibility test. \textit{Journal of the Acoustical Society of
America} 96 (4): 2330--2342.

\item [Munk]W., P. Worcester, and C. Wunsch. 1995. \textit{Ocean Acoustic
Tomography}.  Cambridge: University Press.

\item [Munk]W. and C. Wunsch 1998. Abyssal recipes II. \textit{Deep-Sea Research}
45: 1976--2009.

\item [NAS]National Academy of Sciences. 1963. \textit{Ocean Wave Spectra: Proceedings of a Conference}.
Englewood Cliffs, New Jersey: Prentice-Hall.

\item [Neal]V.T., S. Neshyba, and W. Denner. 1969. Thermal
stratification in the Arctic Ocean. \textit{Science} 166 (3903): 373--374.

\item[Nerem]R.S., E. Leuliette, and A. Cazenave. 2006. Present-day sea-level change: A review. \textit{Comptes Rendus Geosciences} 338 (14--15): 1077--1083.
	
\item [Neumann]G., and W.J. Pierson.
1966. \textit{Principles of Physical Oceanography}.  New Jersey: Prentice-Hall.

\item [Newton]P. 1999. A manual for planetary management. \textit{Nature} 400
(6743): 399.

\item [Niiler]P.P., R.E. Davis, and H.J. White. 1987. Water-following
characteristics of a mix\-ed layer drifter. \textit{Deep-Sea Research} 34 (11):
1867--1881.

\item [Niiler]P.P., A.S. Sybrandy, K. Bi, P.M. Poulain, and D. Bitterman. 1995.
Measurement of the water following capability of holey-sock and \textsc{tristar} drifters. \textit{Deep-Sea Research} 42 (11/12): 1951--1964.

\item [North]G.R. and S. Nakamoto. 1989. Formalism for comparing rain estimation designs. \textit{Journal of Atmospheric and Oceanic Technology}. 6 (6): 985--992.

\item [Oberhuber]J.M. 1988. An atlas based on the \textsc{coads} data set:
The budgets of heat, buoyancy and turbulent kinetic energy at the surface of the
global ocean. \textit{Max-Planck-Institut f\"{u}r Meteorologie}: Report 15.

\item [Open University]1989a. \textit{Ocean Circulation}.  Oxford: Pergamon
Press.

\item[Open University]1989b. \textit{Seawater: Its Composition, Properties and
Behaviour}.  Oxford: Pergamon Press.

\item [Open University]1989c. \textit{Waves, Tides and Shallow
Water--Processes}.  Oxford: Pergamon Press.

\item [Oppenheim]A.V. and R.W. Schafer. 1975. \textit{Digital Signal Processing}. 
Englewood Cliffs, New Jersey: Prentice-Hall.

\item [Orsi]A.H., T. Whitworth and W.D. Nowlin. 1995. On the meridional
extent and fronts of the Antarctic Circumpolar Current. \textit{Deep-Sea Research}
42(5): 641--673.

\item [Orsi]A.H., G.C. Johnson, and J.L. Bullister. 1999. Circulation,
mixing, and production of Antarctic Bottom Water.
\textit{Progress in Oceanography} 43 55-109.

\item [Pacanowski]R., and S.G.H. Philander. 1981. Parameterization of vertical
mixing in numerical models of tropical oceans. \textit{Journal of Physical
Oceanography} 11: 1443--1451.

\item [Pacanowski]R.C., and S.M. Griffies. 1999 \textit{MOM 3.0 Manual}.
\textsc{Noaa}/Geophysical Fluid Dynamics Laboratory, Princeton.

\item [Parke]M.E., R.H. Stewart, D.L. Farless, and D.E. Cartwright. 1987. On
the choice of orbits for an altimetric satellite to study ocean circulation and
tides. \textit{Journal of Geophysical Research} 92: 11,693--11,707.

\item [Parker]R.L. 1994. \textit{Geophysical Inverse Theory}. Princeton:
Princeton University Press.

\item [Pedlosky]J. 1987. \textit{Geophysical Fluid Dynamics}. 2nd ed.
Berlin: Springer Verlag.

\item [Pedlosky]J. 1996. \textit{Ocean Circulation Theory}.  Berlin:
Springer--Verlag.

\item [Percival]D.B., and A.T. Walden. 1993. \textit{Spectral Analysis for
Physical Applications: Multitaper and Conventional Univariate Techniques}. 
Cambridge: University Press.

\item [Philander]S.G.H., T. Yamagata, and R.C. Pacanowski. 1984. Unstable
air-sea interactions in the tropics. \textit{Journal of Atmospheric Research} 41:
604--613.

\item [Philander]S.G. 1990. \textit{El Ni\~{n}o, La Ni\~{n}a, and the Southern
Oscillation}. Academic Press.

\item [Phillips]O.M. 1957. On the generation of waves by turbulent wind.
\textit{Journal of Fluid Mechanics}. 2 (5): 417--445.

\item [Phillips]O.M. 1960. On the dynamics of unsteady gravity waves of finite
amplitude.  Part I.  The elementary interactions. \textit{Journal of Fluid
Mechanics} 9 (2): 193--217.

\item [Picaut]J., F. Masia, and Y.d. Penhoat. 1997. An advective-reflective
conceptual model for the oscillatory nature of the ENSO. \textit{Science} 277
(5326): 663--666.

\item [Pickard]G.L., and W.J. Emery. 1990. \textit{Descriptive Physical
Oceanography: An Introduction}. 5th enlarged ed. Oxford: Pergamon Press.

\item [Pierson]W.J., and L. Moskowitz. 1964. A proposed spectral form for fully
developed wind seas based on the similarity theory of S.A. Kitaigordskii.
\textit{Journal of Geophysical. Research} 69: 5181--5190.

\item [Pinet]P.R. 2000. \textit{Invitation to oceanography}. 2nd Edition.
Sudbury, Massachusetts: Jones and Bartlett Publishers.

\item [Polzin]K.L., J.M. Toole, J.R. Ledwell, and R.W. Schmitt. 1997. Spatial
variability of turbulent mixing in the abyssal ocean. \textit{Science} 276
(5309): 93--96.

\item[Post]D.E. and L.G. Votta. 2005. Computational science demands a new paradigm. \textit{Physics Today} 58 (1): 35--41.
	
\item [Powell]M. D., P.J. Vickery and T.A. Reinhold. 2003. Reduced drag
coefficient for high wind speeds in tropical cyclones. \textit{Nature} 422 (6929):
279--283.

\item [Press]W.H., S.A. Teukolsky, W.T. Vetterling, and B.P. Flannery. 1992.
\textit{Numerical Recipes in FORTRAN}. 2nd ed.  Cambridge: University Press.

\item [Preston-Thomas]H. 1990. The International Temperature Scale of 1990
(\textsc{its}-90). \textit{Metrologia} 27 (1): 3--10.

\item [Proudman]J. 1916. On the motion of solids in a liquid possessing
vorticity. \textit{Proceedings Royal Society (London)} A 92: 408--424 

\item [Pugh]D.T. 1987. \textit{Tides, Surges, and Mean Sea-Level}.  Chichester:
John Wiley.

\item [Rahmstorf]S. 1995. Bifurcations of the Atlantic thermohaline circulation
in response to changes in the hydrological cycle. \textit{Nature} 378: 145--149.

\item [Ralph]E.A., and P.P. Niiler. 2000. Wind-driven currents in the tropical
Pacific. \textit{Journal of Physical Oceanography} 29 (9): 2121--2129.

\item [Ramanathan]V., B. Subasilar, G.J. Zhang, W. Conant, R.D. Cess,
J.T. Kiehl, H. Grassl, and L. Shi. 1995. Warm pool heat budget shortwave cloud
forcing: A missing physics? \textit{Science} 267 (5197): 499--503.

\item [Rasmusson]E.M., and T.H. Carpenter. 1982. Variations in tropical sea
surface temperature and surface wind fields associated with the Southern
Oscillation/El Ni\~{n}o. \textit{Monthly Weather Review} 110: 354--384.

\item [Rasmusson]E.M., and J.M. Wallace. 1983. Meteorological Aspects of the
El Ni\~{n}o/Southern Oscillation. \textit{Science} 222: 1195--1202.

\item [Reid]R.O. 1948. The equatorial currents of the eastern Pacific as
maintained by the stress of the wind. \textit{Journal of Marine Research} 7 (2):
75--99.

\item [Reynolds]O. 1883. An experimental investigation of the circumstances
which determine whether the motion of water will be direct or sinuous, and the
law of resistance in parallel channels. \textit{Philosophical Transactions, Royal
Society London} 174: 935. 

\item [Reynolds]R.W., and D.C. Marsico. 1993. An improved real-time global sea
surface temperature analysis. \textit{Journal of Climate} 6: 114--119.

\item [Reynolds]R.W., and T.M. Smith. 1994. Improved global sea surface
temperature analysis using optimum interpolation. \textit{Journal of Climate} 7:
929--948.

\item [Reynolds]R.W., and T.M. Smith. 1995. A high-resolution global sea surface 
climatology. \textit{Journal of Climate} 8 (6): 1571--1583.

\item [Reynolds]R.W., N.A. Rayner, T. M. Smith, D. C. Stokes, W. Wang. 2002. An improved in situ and satellite SST analysis for climate. \textit{Journal of Climate} 15: 1609--1625.

\item [Rhines]P.B. 1984. Notes on the general circulation of the ocean. In:
\textit{A Celebration in Geophysics and Oceanography--1982}. 83--86. Scripps
Institution of Oceanography.

\item [Richardson]E.G. 1961. \textit{Dynamics of Real Fluids}. 2nd ed. London:
Edward Arnolds.

\item [Richardson]P.L. 1981. Eddy kinetic energy in the North Atlantic
from surface drifters. \textit{Journal of Geophysical Research}. 88 (C7): 4355--4367.

\item [Richardson]P.L. 1993. Tracking Ocean Eddies. \textit{American Scientist}
81: 261--271.

\item [Richardson]P. L. 2008. On the history of meridional overturning circulation schematic diagrams. \textit{Progress In Oceanography} 76 (4): 466--486.

\item [Ring Group]R.H. Backus, G.R. Flierl, D.R. Kester, D.B. Olson, P.L. Richardson, A.C. Vastano, P.H. Wiebe, and J.H. Wormuth. 1981. Gulf Stream cold-core rings: Their physics, chemistry, and
biology. \textit{Science} 212 (4499): 1091--1100.

\item [Rintoul]S.R., and C. Wunsch. 1991. Mass, heat, oxygen and nutrient
fluxes and budgets in the North Atlantic Ocean. \textit{Deep-Sea Research} 38
(Supplement 1): S355--S377.

\item [Riser]S. C., L. Ren, et al. 2008. Salinity in ARGO. \textit{Oceanography} 21 (1): 56--67.

\item [Robinson]A.R., S.M. Glenn, M.A. Spall, L.J. Walstad, G.M. Gardner, and
W.G. Leslie. 1989. Forecasting Gulf Stream meanders and rings. \textit{EOS
Transactions American Geophysical Union} 70: (45).

\item [Roed]L.P., B. Hackett, B. Gjevik, and L.I. Eide. 1995. A review of the Metocean Modeling Project (\textsc{mompop}) Part 1: Model comparison study. In: \textit{Quantitative Skill Assessment for Coastal Ocean Models}. Edited by D. R. Lynch and A. M. Davies. 285--305. Washington DC: American Geophysical Union.

\item [Ropelewski]C.F., and M.S. Halpert. 1987. Global and regional precipitation associated with El Ni\~{n}o/Southern Oscillation. \textit{Monthly Weather Review} 115: 1606--1626.

\item [Rossby]C.C. 1936. Dynamics of steady ocean currents in the light of experimental fluid mechanics. \textit{Papers in Physical Oceanography and Meteorology, Massachusetts Institute of Technology and Woods Hole Oceanographic Institution}. 5(1): 43.

\item [Rossow]W.B. and R.A. Schiffer. 1991. \textsc{Isccp} Cloud Data Products. \textit{Bulletin of the  American Meteorology Society}, 72 (1): 2--20.

\item [Rudnick], D.L., T.J. Boyd, R.E. Brainard, G.S. Carter, G.D. Egbert, M.C. Gregg, P.E. Holloway, J.M. Klymak, E. Kunze, C.M. Lee, M.D. Levine, D.S. Luther, J.P. Martin, M.A. Merrifield, J.N. Moum, J.D. Nash, R. Pinkel, L. Rainville, T.B. Sanford. 2003. From Tides to Mixing Along the Hawaiian Ridge. \textit{Science}  301 (5631): 355--357.

\item [Sabine]C.L., R.A. Feely, N. Gruber, R.M. Key, K. Lee, J.L. Bullister, R. Wanninkhof, C.S. Wong, D.W.R. Wallace, B. Tilbrook, F.J. Millero, T.-H. Peng, A. Kozyr, T. Ono and A.F. Rios. 2004. The Oceanic Sink for Anthropogenic CO$_2$. \textit{Science} 305 (5682): 367--371.

\item [Sandwell]D.T., and W.H.F. Smith. 2001. Bathymetric Estimation. In: \textit{Satellite Altimetry and Earth Sciences}. Edited by L.-L. Fu and A. Cazanave. 441--457. San Diego: Academic Press.

\item [Satake]K., K. Shimazaki, Y. Tsuji, and K. Ueda. 1996. Time and size of a
giant earthquake in Cascadia inferred from Japanese tsunami records of January
1700. \textit{Nature} 379 (6562): 246--249.

\item [Saunders]P.M. and N. P. Fofonoff. 1976. Conversion of pressure to
depth in the ocean. \textit{Deep-Sea Research} 23: 109--111.

\item [Saunders]P.M. 1986. The accuracy of measurements of salinity, oxygen and
temperature in the deep ocean. \textit{Journal of Physical Oceanography} 16 (1): 189--195.

\item [Schmitt]R.W., H. Perkins, J.D. Boyd, and M.C. Stalcup. 1987. C-SALT: An
investigation of the thermohaline staircase in the western tropical North
Atlantic. \textit{Deep-Sea Research} 34 (10): 1655--1665.

\item [Schmitt]R.W., P.S. Bogden, and C.E. Dorman. 1989. Evaporation
minus precipitation and density fluxes for the North Atlantic. \textit{Journal of Physical Oceanography} 19: 1208--1221.

\item [Schmitt]R.W. 1994. The ocean freshwater cycle. \textsc{jsc} Ocean Observing
System Development Panel, Texas A\&M University, College Station, Texas: 40.

\item [Schmitz]W.J. 1996. On the World Ocean Circulation: Volume I. Some Global
Features/ North Atlantic Circulation. Woods Hole Oceanographic Institution,
Technical Report \textsc{whoi}--96--03.

\item [Schubert]S.D., R.B. Rood, and J. Pfaendtner. 1993. An assimilated
dataset for Earth science applications. \textit{Bulletin American Meteorological
Society} 74 (12): 2331--2342.

\item [Selby]J.E.A., and R.A. McClatchey. 1975. Atmospheric transmittance from 0.25 to 28.5 $\mu$m: Computer code \textsc{lowtran} 3. Air Force Cambridge Research Laboratories, Optical Physics Laboratory Technical Report TR--75--0255.

\item [Service]R.F. 1996. Rock chemistry traces ancient traders. \textit{Science} 274 (5295): 2012--2013.

\item[Sette]O.E., and J.D. Isaacs. 1960. Symposium on ``The Changing Pacific Ocean in 1957 and 1958''. California Cooperative Oceanic Fisheries Investigations Reports VII: 13--217.

\item [Shamos]M.H. 1995. \textit{The Myth of Scientific Literacy}. New Brunswick:
Rutger  University Press.

\item [Shchepetkin]A. F. and J.C. McWilliams. 2004. The Regional Oceanic Modeling System: A split-explicit, free-surface, topography-following-coordinate ocean model. \textit{Ocean Modelling} 9: 347--404.

\item [Shepard]F.P., G.A. MacDonald and D.C. Cox. 1950. The tsunami of April 1, 1946. \textit{Bulletin of the Scripps Institution of Oceanography} 5(6): 391--528.

\item [Shepard]F.P. 1963. \textit{Submarine Geology}. 2nd ed. New York: Harper and Row.

\item [Slingo]J.M., K.R. Sperber, J.S. Boyle, J.-P. Ceron, M. Dix, B. Dugas, W. Ebisuzaki, J. Fyfe, D. Gregory, J.-F. Gueremy, J. Hack, A. Harzallah, P. Inness, A. Kitoh, W.K.-M. Lau, B. McAvaney, R. Madden, A. Matthews, T.N. Palmer, C.-K. Park, D. Randall, and N. Renno. 1995. Atmospheric Model Intercomparison Project (\textsc{amip}): Intraseasonal Oscillations in 15 Atmospheric General Circulation Models (Results From an \textsc{aimp} Diagnostics Subproject). World Meteorological Organization/World Climate Research Programme, WCRP--88 (WMO/TD No. 661).

\item [Slutz]R.J., S.J. Lubker, J.D. Hiscox, S.D. Woodruff, R.L. Jenne, D.H. Joseph, P.M. Steurer, and J.D. Elms. 1985: Comprehensive Ocean-Atmosphere Data Set; Release 1. \textsc{noaa} Environmental Research Laboratories, Climate Research Program, Boulder, Colorado:, 268. (NTIS PB86--105723).

\item [Smagorinski]J. 1963. General circulation experiments with primitive
equations I. The basic experiment. \textit{Monthly Weather Review} 91: 99--164.

\item [Smith]R.D., M.E. Maltrud, F.O. Bryan and M.W. Hecht. 2000. Numerical simulation of the North Atlantic ocean at 1/10\degrees. \textit{Journal of Physical Oceanography} 30 (7): 1532--1561.

\item [Smith]S.D. 1980. Wind stress and heat flux over the ocean in gale force winds. \textit{Journal of Physical Oceanography} 10: 709-726.
	
\item [Smith]S.D. 1988. Coefficients for sea surface wind stress, heat flux and wind profiles as a function of wind speed and temperature. \textit{Journal of Geophysical Research} 93: 15,467--15,472.

\item [Smith]T.M., T.R. Karl and R.W. Reynolds. 2002. How accurate are climate simulations? \textit{Science} 296 (5567): 483--484.

\item [Smith]T.M., and R.W. Reynolds. 2004. Improved extended reconstruction of \textsc{sst} (1854--1997). \textit{Journal of Climate} 17 (6): 2466--2477.
	
\item [Smith]W.H.F., and D.T. Sandwell. 1994. Bathymetric prediction from dense satellite altimetry and sparse shipboard bathymetry. \textit{Journal of Geophysical Research} 99 (B11): 21,803--21,824.

\item [Smith]W.H.F. and D.T. Sandwell 1997. Global sea floor topography from satellite altimetry and ship depth soundings. \textit{Science} 277 (5334): 1956--1962.

\item[Snodgrass]F.E. 1964. Precision digital tide gauge. \textit{Science} 146 (3641): 198--208.

\item [Soulen]R.J., and W.E. Fogle. 1997. Temperature scales below 1 kelvin. \textit{Physics Today} 50 (8 Part 1): 36--42.

\item [Stammer]D., R. Tokmakian, A. Semtner, and C. Wunsch. 1996. How well does a 1/4\degrees\ global circulation model simulate large-scale oceanic observations? \textit{Journal of Geophysical Research} 101 (C10): 25,779--25,811.

\item [Starr]V.P. 1968. \textit{Physics of Negative Viscosity Phenomena}.  New York: McGraw-Hill.

\item [Steig]E.J. 2006. Climate change: The south-north connection. \textit{Nature} 444 (7116): 152--153.

\item [Stern]M.E. 1960. The `salt fountain' and thermohaline convection. \textit{Tellus} 12: 172--175.

\item [Stewart]I. 1992. Warning---handle with care! \textit{Nature} \textbf{355}:
16--17.

\item [Stewart]R.H. 1980. Ocean wave measurement techniques. In: \textit{Air Sea
Interaction, Instruments and Methods}. Edited by L. H. F. Dobson and R. Davis.
447--470. New York: Plenum Press.

\item [Stewart]R.H. 1985. \textit{Methods of Satellite Oceanography}. University
of California Press.

\item [Stewart]R.H. 1995. Predictability of Texas Rainfall Patterns on Time
Scales of Six to Twelve Months: A Review. In: \textit{The Changing Climate of
Texas: Predictability and Implications for the Future} Edited by J. Norwine, J.R.
Giardino, G.R. North and J.B. Valdes. College Station, Texas: Texas A\&M
University. 38--47

\item [Stocker]T.F., and O. Marchal. 2000. Abrupt climate change in the
computer: Is it real? \textit{Proceedings National Academy of Sciences} 97 (4):
1362--1365.

\item [Stokes]G.G. 1847. On the theory of oscillatory waves. \textit{Cambridge Transactions} 8: 441--473.

\item [Stommel]H. 1948. The westward intensification of wind-driven ocean
currents. \textit{Transactions, American Geophysical Union} 29 (2): 202--206.

\item[Stommel]H. 1958. The abyssal circulation. \textit{Deep-Sea Research} 5 (1):
80--82.

\item [Stommel]H., A.B. Arons, and A.J. Faller. 1958. Some examples of
stationary flow patterns in bounded basins. \textit{Tellus} 10 (2): 179--187.

\item [Stommel]H., and A.B. Arons. 1960. On the abyssal circulation of the
world ocean---II. An idealized model of the circulation pattern and amplitude in
oceanic basins. \textit{Deep-Sea Research} 6: 217--233.

\item [Stommel]H.M., and D.W. Moore. 1989. \textit{An Introduction to the Coriolis Force}.  Cambridge: University Press.

\item [Stott]P.A., S.F.B. Tett, G.S. Jones, M.R. Allen, J.F.B. Mitchell and G.J. Jenkins 2000. External Control of 20th Century Temperature by Natural and Anthropogenic Forcings. \textit{Science} 290  (5499): 2133-2137.

\item [Strub]P.T., T.K. Chereskin, P.P. Niiler, C. James, and M.D. Levine.
1997. Altimeter-derived variability of surface velocities in the California
Current System 1. Evaluation of \textsc{Topex} altimeter velocity resolution.
\textit{Journal of Geophysical Research} 102 (C6): 12,727--12,748.

\item [SUN]Working Group on Symbols, Units and Nomenclature in Physical Oceanography. 1985. \textit{The International System of units (SI) in oceanography}. \textsc{iapso}  Paris: \textsc{Unesco} Technical Papers in Marine Science 45: 124.

\item [Sverdrup]H.U. 1947. Wind-driven currents in a baroclinic ocean: with
application to the equatorial currents of the eastern Pacific.
\textit{Proceedings of the National Academy of Sciences} 33 (11): 318--326.

\item [Sverdrup]H.U., M.W. Johnson, and R.H. Fleming. 1942. \textit{The Oceans:
Their Physics, Chemistry, and General Biology}.  Englewood Cliffs, New Jersey:
Prentice-Hall.

\item [SWAMP Group]Sea Wave Modeling Project. 1985. \textit{Ocean Wave Modeling}. New York: Plenum Press.

\item [Swenson]K.R., and A.E. Shaw. 1990. The Argos system: Monitoring the world's environment. \textit{Oceanography} 3 (1): 60--61.

\item [Takayabu]Y.N., T. Ihuchi, M. Kachi, A. Shibata, and H. Kanzawa. 1999. Abrupt termination of the 1997--98 El Nino in response to a Madden-Julian oscillation. \textit{Nature} 402 (6759): 279--282.

\item [Tapley]B.D., and M.-C. Kim. 2001. Applications to geodesy. In
\textit{Satellite Altimetry and Earth Sciences}. Edited by L.-L. Fu and A.
Cazenave. 371--406. San Diego: Academic Press.

\item [Taylor]G.I. 1921. Experiments with rotating fluids. \textit{Proceedings
Royal Society (London)} A 100: 114--121.

\item [Taylor]P.K. Editor. 2000. Intercomparison and validation of ocean-atmosphere energy flux fields: Final Report of the Joint World Climate Research Program and Scientific Committee on Ocean Research Working Group on Air-sea Fluxes, \textit{World Climate Research Program Report} \textsc{wcrp}-112:
303.
	
\item [Taylor]P. K., E.F. Bradley, C.W. Fairall, D. Legler, J. Schultz, R.A. Weller and G.H. White. 2001. Surface fluxes and surface reference sites. In: \textit{Observing the Oceans in the 21st Century}. Edited by C.J. Koblinsky and N.R. Smith. Melbourne, Bureau of Meteorology: 177--197.

\item [Tchernia]P. 1980. \textit{Descriptive Regional Oceanography}. Oxford: Pergamon Press.

\item [Tennekes]H., and J.L. Lumley. 1990. \textit{A First Course in Turbulence}.  Boston: MIT Press.

\item [Thurman]H.V. 1985. \textit{Introductory Oceanography}. Fourth ed. Columbus: Charles E. Merrill Publishing Company.

\item [Titov]V.V. and F.I. Gonzalez. 1997. Implementation and testing of the Method of Splitting Tsunami (\textsc{most}) model. \textsc{Noaa} Pacific Marine Environmental Laboratory Contribution 1927: 14.

\item [Toggweiler]J.R. 1994. The ocean's overturning circulation. \textit{Physics Today} 47 (11): 45--50.

\item [Toggweiler]J. R. and J. Russell 2008. Ocean circulation in a warming climate. \textit{Nature} 451 (7176): 286--288.
	
\item [Tolmazin]D. 1985. \textit{Elements of Dynamic Oceanography}. Boston: Allen and Unwin.

\item [Tomczak]M., and J.S. Godfrey. 1994. \textit{Regional Oceanography: An Introduction}. London: Pergamon.

\item [Tomczak]M. 1999. Some historical, theoretical and applied aspects of quantitative water mass analysis. \textit{Journal of Marine Research} 57 (2): 275--303.

\item [Trenberth]K.E., and D.J. Shea. 1987. On the evolution of the Southern Oscillation. \textit{Monthly Weather Review} 115: 3078--3096.

\item [Trenberth]K.E., W.G. Large, and J.G. Olson. 1989. The effective drag coefficient for evaluating wind stress over the oceans. \textit{Journal of Climate}, 2:
1507--1516.

\item [Trenberth]K.E., W.G. Large, and J.G. Olson. 1990. The mean annual cycle
in global ocean wind stress. \textit{Journal of Physical Oceanography} 20 (11):
1742--1760.

\item [Trenberth]K.E., and A. Solomon. 1994. The global heat balance: heat
transports in the atmosphere and ocean. \textit{Climate Dynamics} 10 (3):
107--134.

\item [Trenberth]K.E. 1997. The use and abuse of climate models. \textit{Nature} 386 (6621): 131--133.

\item [Trenberth]K.E. 1997. The definition of El Ni\~{n}no. \textit{Bulletin of the American Meteorological Society} 78 (12): 2771--2777.

\item [Trenberth]K.E. and J.M. Caron. 2001. Estimates of meridional atmospheric and oceanic heat transports. \textit{Journal of Climate} 14 (16): 3433--3443.

\item [Uchida]H., S. Imawaki, and J.-H. Hu. 1998. Comparisons of Kuroshio surface velocities
derived from satellite altimeter and drifting buoy data. \textit{Journal of Oceanography} 54: 115--122.

\item [Uppala]S.M., and P.W. K\r{a}llberg, A.J. Simmons, U. Andrae, V. Da Costa Bechtold, M. Fiorino, J. K. Gibson, J. Haseler, A. Hernandez, G. A. Kelly, X. Li, K. Onogi, S. Saarinen, N. Sokka, R. P. Allan, E. Andersson, K. Arpe, M. A. Balmaseda, A. C. M. Beljaars, L. Van De Berg, J. Bidlot, N. Bormann, S. Caires, F. Chevallier, A. Dethof, M. Dragosavac, M. Fisher, M. Fuentes, S. Hagemann, E. H\'{o}lm, B. J. Hoskins, L. Isaksen, P.A.E.M. Janssen, R. Jenne, A. P. McNally, J.-F. Mahfouf, J.-J. Morcrette, N. A. Rayner, R. W. Saunders, P. Simon, A. Sterl, K. E. Trenberth, A. Untch, D. Vasiljevic, P. Viterbo, J. Woollen. 2005. The \textsc{era}--40 re-analysis. \textit{Quarterly Journal of the Royal Meteorological Society} 131(612): 2961--3012.

\item [Ursell]F. 1950. On the theoretical form of ocean swell on a rotating
earth. \textit{Monthly Notices Royal Astronomical Society, Geophysical Supplement} 6 (1):
G1--G8.

\item [van Meurs]P. 1998. Interactions between near-inertial mixed layer
currents and the meso\-scale: The importance of spatial variability in the
vorticity field. \textit{Journal of Physical Oceanography} 28 (7): 1363--1388.

\item [Vesecky]J.F., and R.H. Stewart. 1982. The observation of ocean surface
phenomena using imagery from the Seasat synthetic aperture radar: An assessment.
\textit{Journal of Geophysical Research} 87 (C5): 3397--3430.

\item [von Arx]W.S. 1962. \textit{An Introduction  to Physical Oceanography}. Reading, Massachusetts: Addison-Wesley.

\item [WAMDI Group]S. Hasselmann, K. Hasselmann, E. Bauer, P.A.E.M. Jans\-sen,
G.J. Komen, L. Bertotti, P. Lionello, A. Guillaume, V.C. Cardone, J.A. Greenwood,
M. Reistad, L. Zambresky, and J.A. Ewing. 1988. The WAM model---A third
generation wave prediction model. \textit{Journal of Physical Oceanography} 18:
1775--1810.

\item [Warren]B.A. 1973. Transpacific hydrographic sections at Latitudes 43\degrees\ S and 28\degrees\ S: The \textsc{scorpio} Expedition--II. Deep Water. \textit{Deep-Sea Research} 20: 9--38.
	
\item [WCRP]World Climate Research Program. 1995. \textit{Proceedings of the Workshop on Global Coupled General Circulation Models}. World Meteorological Organization/World Climate Research
Program, WCRP-87 (WMO/TD Number 655).

\item [Weaver]A.J. and C. Hillaire-Marcel 2004. Global Warming and the Next Ice Age. \textit{Science} 304 (5669): 400--402.

\item [Webb]D.J., and N. Suginohara. 2001. Vertical mixing in the ocean.
\textit{Nature} 409 (6816): 37.

\item [Webster]F. 1968. Observations of inertial-period motions in the deep
sea. Reviews of Geophysics 6 (4): 473--490.

\item [Webster]P.J., and R. Lukas. 1992. \textsc{toga coare}: The Coupled
Ocean-Atmosphere Response Experiment. \textit{Bulletin American Meteorological
Society} 73 (9): 1377--1416.

\item [Weller]R.A., J.P. Dean, J. Marra, J.F. Price, E.A. Francis, and D.C.
Boardman. 1985. Three--dimensional flow in the upper ocean. \textit{Science} 227:
1552--1556.

\item [Weller]R.A., and A.J. Plueddmann. 1996. Observations of the vertical
structure of the oceanic boundary layer. \textit{Journal of Geophysical Research}
101 (C4): 8,789--8,806.

\item [Wentz]P.J., S. Peteherych, and L.A. Thomas. 1984. A model function for
ocean radar cross sections at 14.6 GHz. \textit{Journal of Geophysical Research}
89 (C3): 3689--3704.

\item [West]G.B. 1982. Mean Earth ellipsoid determined from Seasat 1 altimetric
observations. \textit{Journal of Geophysical Research} 87 (B7): 5538--5540.

\item [White]G. Editor. 1996. \textsc{wcrp} Workshop on Air-Sea Flux Fields for Forcing Ocean Models and Validating GCMs. Geneva: World Meteorological Organization Report WCRP--95 (WMO/TD Number 762).

\item [White]W.B., and D.R. Cayan. 1998. Quasi-periodicity and global symmetries in interdecadal upper ocean temperature variability. \textit{Journal of Geophysical Research} 103 (C10): 21,335--21,354.

\item [Whitham]G.B. 1974. \textit{Linear and Nonlinear Waves}.  New York: John Wiley.

\item [Whittaker]E.T., and G.N. Watson. 1963. \textit{A Course of Modern Analysis}. 4th ed.  Cambridge: University Press.

\item [Whitworth]T., and R.G. Peterson. 1985. Volume transport of the Antarctic
Circumpolar Current from bottom pressure measurements. \textit{Journal of Physical
Oceanography} 15 (6): 810--816.

\item [Wiegel]R.L. 1964. \textit{Oceanographical Engineering}.  Englewood Cliffs, New Jersey: Prentice Hall.

\item [Wilson]W.D. 1960. Equation for the speed of sound in sea water. \textit{Journal of the Acoustical Society of America} 32 (10): 1357.

\item [Woodruff]S.D., R.J. Slutz, R.L. Jenne, and P.M. Steurer. 1987. A comprehensive ocean--atmosphere data set. \textit{Bulletin American Meteorological Society} 68: 1239--1250.

\item [Wooster]W.S. 1960. Investigations of equatorial undercurrents. \textit{Deep-Sea Research} 6 (4): 263--264.

\item [Wooster]W.S., A.J. Lee, and G. Dietrich. 1969. Redefinition of salinity.
\textit{Deep-Sea Research} 16 (3): 321--322.

\item [Worthington]L.V. 1981. The water masses of the World Ocean: Some results of a fine-scale census. In: \textit{Evolution of Physical Oceanography: Scientific surveys in honor of Henry Stommel}. Edited by B. A. Warren and C. Wunsch. 42--69. Cambridge: Massachusetts Institute of Technology.

\item [Wunsch]C. 1996. \textit{The Ocean Circulation Inverse Problem}.  Cambridge: University Press.

\item [Wunsch]C. 2002. Ocean observations and the climate forecast problem.In: \textit{Meteorology at the Millennium}, Edited by R.P. Pearce. London: Royal Meteorological Society: 233--245.

\item [Wunsch]C. 2002b. What is the thermohaline circulation? \textit{Science} 298 (5596): 1179--1180.

\item [Wust]G. 1964. The major deep--sea expeditions and research vessels 1873--1960. \textit{Progress in Oceanography} 2: 3--52.

\item [Wyrtki]K. 1975. El Ni\~{n}o---The dynamic response of the equatorial
Pacific Ocean to atmospheric forcing. \textit{Journal of Physical Oceanography} 5
(4): 572--584.

\item [Wyrtki]K. 1979. Sea level variations: monitoring the breath of the Pacific. \textit{EOS Transactions of the American Geophysical Union} 60 (3): 25-27.

\item [Wyrtki]K. 1985. Water displacements in the Pacific and the genesis of El Ni\~{n}o cycles. \textit{Journal of Geophysical Research} 90 (C4): 7129-7132.
	
\item [Xie]P., and P.A. Arkin. 1997. Global precipitation: A 17-year monthly analysis based on gauge observations, satellite estimate, and numerical model outputs. \textit{ Bulletin of the American Meteorological Society} 78 (1):
2539--2558.

\item [Yelland]M., and P.K. Taylor. 1996. Wind stress measurements from the open
ocean. \textit{Journal of Physical Oceanography} 26 (4): 541--558.

\item [Yelland]M.J., B.I. Moat, P.K. Taylor, R.W. Pascal, J. Hutchings and V.C. Cornell. 1998. Wind stress measurements from the open ocean corrected for airflow distortion by the ship. \textit{Journal of Physical Oceanography} 28 (7): 1511--1526.

\item [Yu]X., and M.J. McPhaden. 1999. Dynamical analysis of seasonal and interannual variability in the equatorial Pacific. \textit{Journal of Physical Oceanography} 29 (9): 2350--2369.

\item [Zahn]R. 1994. Core correlations. \textit{Nature} 371 (6495): 289--290.

\end{description}